Supplementary Material for ``Provisional Title''

### 2021-04-15 draft/supp.tex
more supp methods added

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{mathtools} % nice matrices
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage{xcolor} % text color
\usepackage[margin=3cm]{geometry} % margins

@@

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\renewcommand{\thefigure}{\textbf{S\arabic{figure}}} % preceed the Suppl Mat figures with S
\renewcommand{\thetable}{\textbf{S\arabic{table}}} % preceed the Suppl Mat tables with S
\renewcommand{\theequation}{S\arabic{equation}} % preceed the Suppl Mat equations with S
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Fig.}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions

% title, authors, affiliations
\title{Supplementary Material for ``Provisional Title'' }
\author[1,2]{Authors}
\author[1,$\dagger$]{Álvaro Sánchez}
\affil[1]{Department of Ecology \& Evolutionary Biology and Microbial
  Sciences Institute, Yale University, New Haven, CT, USA}
\affil[2]{Other affiliations...}
\affil[$\dagger$]{To whom correspondence should be addressed: \normalfont alvaro.sanchez@yale.edu}
\date{}

@@

Thresholds for vector similarity in ESV space, etc.

@@

Performing 16S rRNA gene sequencing on such a community yields a list of
sequences (exact sequence variants or ESVs) and their respective abundance.
Normalizing by the total number of sequences, we can obtain a vector $\mathbf{s}
= \left( s_1, s_2, \dotsc \right)$ that also satisfies the normalization
conditions of equations \ref{supp-eqn:normx_1} and \ref{supp-eqn:normx_2}.
Only if there are as many ESVs as species does $\mathbf{s}$ have length $N$
(the same as $\mathbf{x}$), but in the most general case it is possible that
\textit{a)} multiple species share a same ESV and/or \textit{b)} species carrying
multiple copies
of the 16S rRNA gene have different sequences for each copy\mr. We denote
the length of $\mathbf{s}$ as $M$, which need not be equal to $N$.

@@

\begin{equation}
f \left( \mathbf{\hat{x}} \right) =
\left| \boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) \right| ^2 =
\sum_i \varepsilon_i^2 \left( \mathbf{\hat{x}} \right)
\end{equation}
%
We can also define a function $\mathbf{h} \left( \mathbf{\hat{x}} \right)$ as

\begin{equation}
\mathbf{h} \left( \mathbf{\hat{x}} \right) = 1 - \mathbf{\hat{x}}
\end{equation}
%
The best estimation for the species composition of a community will be the
$\mathbf{\hat{x}}$ that minimizes $f \left( \mathbf{\hat{x}} \right)$ while
satisfying the normalization condition (eq. \ref{supp-eqn:normx_1}) and the
boundary conditions (eq. \ref{supp-eqn:normx_2}). We solve
this problem using the \texttt{nloptr} package for R\mr, passing it
$f \left( \mathbf{\hat{x}} \right)$ as the function to minimize,
$\mathbf{h} \left( \mathbf{\hat{x}} \right) = 0$ as an equality constraint,
and 0 and 1
as the lower and upper bounds for the entries of $\mathbf{\hat{x}}$. Further
details can be found in the code for the analysis (see section 
\nameref{supp-datacode} of this Supplementary Material).

@@

% references
%{\small
  %% \bibliography{refs}
%  \bibliography{extracted-supp}
% \bibliographystyle{unsrt}
%% \bibliographystyle{mystyle}
%\bibliographystyle{mybibwithurl}
%}

>>>>

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{mathtools} % nice matrices
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage{xcolor} % text color
\usepackage[margin=3cm]{geometry} % margins
\usepackage[nottoc,numbib]{tocbibind} % add references to table of contents

@@

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\addto{\captionsenglish}{\renewcommand{\bibname}{Supplementary References}}
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\renewcommand{\thefigure}{\textbf{S\arabic{figure}}} % preceed the Suppl Mat figures with S
\renewcommand{\thetable}{\textbf{S\arabic{table}}} % preceed the Suppl Mat tables with S
\renewcommand{\theequation}{S\arabic{equation}} % preceed the Suppl Mat equations with S
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Fig.}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions
\hypersetup{citecolor={blue}}

% title, authors, affiliations
\title{Supplementary Material for ``Provisional Title'' }
\author[1,2,3]{Authors}
\author[1,2,$\dagger$]{Álvaro Sánchez}
\affil[1]{Department of Ecology \& Evolutionary Biology,
Yale University, New Haven, CT, USA}
\affil[2]{Microbial Sciences Institute,
Yale University, New Haven, CT, USA}
\affil[3]{Other affiliations...}
\affil[$\dagger$]{To whom correspondence should be addressed: \normalfont alvaro.sanchez@yale.edu}
\date{}

@@

\subsubsection{Examining community composition}
\label{supp-methods:data-proc:community-composition}

We first analyze the composition of the communities assembled in each carbon
source by examining the relative abundances of the exact sequence variants
(ESVs) obtained from 16S rRNA gene sequencing. We define a vector $\mathbf{s}
= \left( s_1, s_2, \cdots, s_M \right)$ such that $s_i$ represents the relative
abundance
of the $i$-th ESV, being $M$ the total number of unique ESVs. We assembled
a total of 16 communities, 8 in glutamine
and 8 in citrate (see Methods \mr), with 3 biological replicates of each. We
start
by comparing the composition of replicate communities in terms of their ESV
relative abundances. To do this, we measure the distance $d$ between two
communities (with ESV abundances $\mathbf{s}$ and $\mathbf{s}'$ respectively) as

\begin{equation}
d \left( \mathbf{s}, \mathbf{s}' \right) =
\sqrt{\sum_i{\left(s_i - s'_i\right)^2}}
\label{supp-eqn:esvdist}
\end{equation}
%
We compute all the distances across replicates of the same community, obtaining
a set of 48 distances (16 communities $\times$ 3 pairwise distances between the
three replicates) that we will denote as $\left\{d\right\}$

\begin{equation} 
\left\{d\right\} = 
\left\{ d\left(\mathrm{community} \; i \; \mathrm{replicate} \; j,
\mathrm{community} \; i \; \mathrm{replicate} \; j' \right) \;
\text{for all $i$, $j$, $j' \neq j$} \right\}
\end{equation}
%
We then define a threshold $d_T$ as

\begin{equation}
d_T = \mathrm{Q}4 \left( \left\{d\right\} \right) + 
\mathrm{IQR} \left( \left\{d\right\} \right)
\label{supp-eqn:threshold}
\end{equation}
%
Where Q4 and IQR represent the fourth quantile and the interquartile range
respectively. Note that this threshold is analogous to the one used in
standard boxplots to identify outliers.

Finally, we discard the samples that are at a distance larger than $d_T$ from
all other replicates of the same community. These are replicate 2 of community
1 in glutamine and replicate 2 of community 8 in citrate (see Figure S1 \mr).
The remaining replicates were used in further analyses. We took the average
ESV abundance across replicates when indicated.

@@

Performing 16S rRNA gene sequencing on such a community yields a list of
sequences (exact sequence variants or ESVs) and their respective abundance.
Normalizing by the total number of sequences, we can obtain a vector $\mathbf{s}
= \left( s_1, s_2, \dotsc \right)$ that also satisfies the normalization
conditions of equations \ref{supp-eqn:normx_1} and \ref{supp-eqn:normx_2}.
Only if there are as many ESVs as species does $\mathbf{s}$ have length $N$
(the same as $\mathbf{x}$), but in the most general case it is possible that
\textit{a)} multiple species share a same ESV and/or \textit{b)} species carrying
multiple copies
of the 16S rRNA gene have different sequences for each copy \cite{Vetrovsky2013}. We denote
the length of $\mathbf{s}$ as $M$, which need not be equal to $N$.

@@

\begin{equation}
f \left( \mathbf{\hat{x}} \right) =
\left| \boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) \right| ^2 =
\sum_i \varepsilon_i^2 \left( \mathbf{\hat{x}} \right)
\end{equation}
%
We can also define a function $h \left( \mathbf{\hat{x}} \right)$ as

\begin{equation}
h \left( \mathbf{\hat{x}} \right) = 1 - \sum_i\hat{x}_i
\end{equation}
%
The best estimation for the species composition of a community will be the
$\mathbf{\hat{x}}$ that minimizes $f \left( \mathbf{\hat{x}} \right)$ while
satisfying the normalization condition (eq. \ref{supp-eqn:normx_1}) and the
boundary conditions (eq. \ref{supp-eqn:normx_2}). We solve
this problem using the \texttt{nloptr} package for R \cite{nloptr}, passing it
$f \left( \mathbf{\hat{x}} \right)$ as the function to minimize,
$h \left( \mathbf{\hat{x}} \right) = 0$ as an equality constraint,
and 0 and 1
as the lower and upper bounds for the entries of $\mathbf{\hat{x}}$, and using
the augmented Lagrangian algorithm \cite{auglag1,auglag2}. Further
details can be found in the code for the analysis (see section 
\nameref{supp-datacode} of this Supplementary Material).

@@

% references
\bibliographystyle{mystyle}
%\bibliographystyle{unsrt}
\bibliography{refs}
